Comparison of linear beam theories

### 2024-03-16 article.tex
fix

\title{Validity of a model for a cantilever two-dimensional beam}

\author{Rudi du Toit}
\address{Department of Mathematics, University of Pretoria, Pretoria, South Africa}

@@

\begin{abstract}
Beams are wildly used in engineering. The the real world, beams are complex three-dimensional objects. But often, simplified models are used to study the vibration of beams. These simplified models are less complex and computationally less expensive. But the validity of these simplified models is not always clear. In this article, the validity of a two-dimensional model for a cantilever beam is investigated. The two-dimensional model is compared to a three-dimensional model.
\end{abstract}

\begin{keyword}
	Beam \sep Cantilever \sep Two-Dimensional Model \sep Three-Dimensional Model \sep Vibration Problems
\end{keyword}

@@

\section{Introduction}
In the article \cite{LVV09}, the authors investigate the validity of a Timoshenko beam using a two-dimensional model as a reference. The two-dimensional model is a simplified model of the Timoshenko beam. The authors use the Finite Element Method (FEM) to solve the eigenvalue problem for the two models. The results are then compared to determine the validity of the two-dimensional model.

In this article, \cite{LVV09} is extended and the validity of a two-dimensional model is investigated, using a three-dimensional model as a reference. Similar to \cite{LVV09}, the investigation focusses on beam-type problems.

The models are presented in a dimensionless variational form. It is also explained how the two-dimensional model is derived from the three-dimensional model, as a special case of plane stress. The existence and uniqueness of solutions for the two models are briefly discussed.

The concept of modal analysis is then introduced using a general vibration problem. From modal analysis, the formal series solution for the vibration problem is obtained. This formal series solution is a linear combination of the eigenvectors and eigenvalues of the problem. This enables us to compare the solutions of the two models, by only considering a comparison of the eigenvectors and eigenvalues.

Using the Finite Element Method (FEM), a formula for the models are obtained to numerically solve the eigenvalue problem for the two models. The eigenvalues are matched up by comparing the mode shapes, and the eigenvalues are then compared.

\section{Models}
\subsection{Three-dimensional model}
Consider a vector valued displacement function $u \in \Omega \subset R^3$.

@@

\subsubsection*{Constitutive Equation}
For an isotropic material, Hooke's law can be written in terms of Young's modulus $E$ and Poisson's ratio $\nu$ as follows.

@@

\subsection{Dimensionless form} \label{ssec:3D_Model:DimensionlessForm}
Suppose $\ell$ represents some notable dimension (e.g.the length) of the elastic body 
and $G$ the shear modulus of elasticity.

@@

\textbf{Remark} The constants introduced are the same as those in \cite{LVV09}. They simplify the equations and allow for comparisons to other models such as the Timoshenko beam model, although that is not covered in this article.

@@

\subsection{Model problems}\label{ssec:3D_Model}
Divide the boundary of $\Omega$ into two distinct parts, referred to as $\Sigma$ and $\Gamma$. The following will be considered a model problem for a three-dimensional elastic body. 
The following will be considered a model problem for a three-dimensional elastic body (also for the two-dimensional elastic body).
\begin{enumerate}
	\item[] The equations of motion \eqref{eq:3D_Model:EM-D} and \eqref{eq:3D_Model:divT-D} is satisfied in $\Omega$;
	\item[] Hooke's law \eqref{eq:3D_Model:CE-D} is satisfied in $\Omega$;
	\item[] The displacement $u = u_\Sigma$ is specified on
		$\Sigma$; The traction $Tn = t_\Gamma$ is specified on
		$\Gamma$.
\end{enumerate} \label{sym:n}\label{sym:SigmaGamma}

\subsubsection{Problem 3D}
Find a vector valued function $u$, satisfying equations
\eqref{eq:3D_Model:EM-D} to \eqref{eq:3D_Model:CE-D} and the following boundary
conditions:

Boundary Conditions:
\begin{eqnarray*}
	u & = & 0 \quad \textrm{ on } \Sigma\\
	Tn & = & 0 \quad \textrm{ on } \Gamma
\end{eqnarray*} With $n$ the outward normal vector to $\Omega$. This model problem is a cantilever three-dimensional body.

\subsection{Variational form}\label{ssec:3D_Model:VariationalForm}
Define the bilinear forms
\begin{eqnarray}
	b(u,\phi) = \int_\Omega~c_1\textrm{Tr}({\cal E}\Phi)+ c_2\textrm{Tr}({\cal E})
	\textrm{Tr}(\Phi) ~dV \label{eq:3D_Model:Bilinear}
\end{eqnarray}\label{sym:b}
and
\begin{eqnarray}
	c(u,\phi) = \int_\Omega~ (\partial^2_t u) \cdot \phi~dV \label{eq:3D_Model:Bilinear_c}
\end{eqnarray}
with $\displaystyle c_1 = \frac{1}{\gamma(1+\nu)}$ and $\displaystyle c_2 = \frac{\nu}{\gamma(1+\nu)(1-2\nu)}$.

@@

The variational problem of Problem 3DV is given as:
\subsubsection{Problem 3DV}\label{sssec:3D_Model:Problem3D1V}
Find a function $u$ such that for all $t>0$, $u \in T(\Omega)$ and
\begin{align}
	c(u,\phi) = -b(u,\phi) + (Q,\phi), \label{eq:3D_Model:Problem3D1VEq}
\end{align}
for all $\phi \in T(\Omega)$

\subsection{Plane stress}\label{ssec:3D_Model:PlaneStress}
Stresses acting parallel to a single plane are referred to as plane stress. 
Consider the right-hand orthonormal set $\left\{e_1, e_2, e_3\right\}$. Without
loss of generality, assume the stresses act parallel to the $e_1$-$e_2$ plane
i.e. $\sigma_{3i} = 0$ for all $i$. From Hooke's Law \eqref{DM_H_E} the
following equations are obtained: \label{sym:e_i}

@@

This is known as general plane stress.

\subsection{Two-dimensional model} \label{ssec:2D_Model}
For the two-dimensional modek, a special case of general plane stress is considered.

Assume that $\sigma_{3i} = 0$ for $i = 1, 2, 3$; $\partial_3 u_1 = 0$, $\partial_3 u_2 = 0$ ($u_1$ and $u_2$ are functions of $x_1$ and $x_2$), and the strain component $\varepsilon_{33} = 0$. Then using the definition of strain 
\begin{eqnarray}
	\partial_i u_3 = 0 \quad \textrm{ for } i = 1,2,3 \textrm{ on } \Omega.
\end{eqnarray}

@@

\subsection*{Equations of motion and constitutive equations}\label{ssec:2D_Model:EquationsOfMotion+ConstitutiveEquations}
 The following equations of motion and constitutive equations follow from subsections \ref{ssec:3D_Model:DimensionlessForm} and 1.3.2.
\subsubsection{Equation of motion}\label{sssec:2D_Model:EquationOfMotion}
\begin{eqnarray}
	\partial_t^2 u & = & \textrm{div}T + Q, \label{eq:2D_Model:EM}
\end{eqnarray}
where
\begin{eqnarray}
	\textrm{div} T & = &
	\begin{bmatrix}
		\partial_1 \sigma_{11} + \partial_2 \sigma_{12} \\
		\partial_1 \sigma_{21} + \partial_2 \sigma_{22}
	\end{bmatrix}. \label{eq:2D_Model:DivT}
\end{eqnarray}
\subsubsection*{Constitutive equation}\label{sssec:2D_Model:ConstitutiveEquation}
\begin{eqnarray}
	T = \frac{1}{\gamma(1+\nu)}\mathcal{E} + \frac{\nu}{\gamma(1-\nu^2)}\textrm{Tr}(\mathcal{E}) I. \label{eq:2D_Model:CE}
\end{eqnarray}

\subsection{Model problem}\label{ssec:2D_Model:ModelProblem}
Divide the boundary of $\Omega$ into two distinct parts, referred to as $\Sigma$ and $\Gamma$.

\subsubsection*{Problem 2D}\label{sssec:2D_Model:Problem2D1}
Find a vector valued function $u$, satisfying equations \eqref{eq:2D_Model:EM} to \eqref{eq:2D_Model:CE} and the boundary conditions:

Boundary Conditions:
\begin{eqnarray*}
	u & = & 0 \quad \textrm{ on } \Sigma,\\
	Tn & = & 0 \quad \textrm{ on } \Gamma.
\end{eqnarray*} with $n$ the outward unit normal. This model problem is a cantilever two-dimensional body.

\subsection{Variational form}\label{ssec:2D_Model:VariationalForm}
Define the bilinear forms
\begin{eqnarray}
	b(u,\phi) = \int_\Omega~c_1\textrm{Tr}({\cal E}\Phi)+ c_2\textrm{Tr}({\cal E})
	\textrm{Tr}(\Phi) ~dA \label{eq:2D_Model:Bilinear}
\end{eqnarray}
and
\begin{eqnarray}
	c(u,\phi) = \int_\Omega~ (\partial^2_t u) \cdot \phi~dA \label{eq:2D_Model:Bilinear_c}
\end{eqnarray}
with $\displaystyle c_1 = \frac{1}{\gamma(1+\nu)}$ and $\displaystyle c_2 = \frac{\nu}{\gamma(1-\nu^2)}$.
Also define the integral
\begin{eqnarray}
	(f,g) &=& \int_{\Omega} f\cdot g \ dA \label{eq:2D_Model:Bilinear_int}
\end{eqnarray}

\subsubsection{Test functions for Problem 2D}\label{sssec:2D_Model:TestFunctions2D1}
Define the following set of test functions $T(\Omega)$ for Problem 2D:
\begin{eqnarray*}
	T(\Omega) & = & \left\{ \phi \in C^1(\bar{\Omega}) \ | \ \phi = 0 \ \textrm{ on } \ \Gamma \right\}. 
\end{eqnarray*}\label{sym:baromega}

The variational problem of Problem 2D is given as:

\subsubsection{Problem 2DV}\label{sssec:2D_Model:Problem2D1V}
Find a function $u$ such that for all $t>0$, $u \in T(\Omega)$ and 
\begin{align}
	c(u,\phi) = -b(u,\phi) + (Q,\phi) \label{eq:2D_Model:Problem2D1VEq}
\end{align}
for all $\phi \in T(\Omega)$.

\section{Existance and uniqueness of solutions}
Our two models, the three-dimensional modela and the two-dimensional model, concern the vibration of elastic bodies. These type of models have a similar variational form as the wave equation and are called second order hyperbolic type problems. The articles \cite{VV02} and \cite{VS18} provide existence and uniqueness results for these type of problems. In \cite{VV02}, it is assumed that $b$ is symmetric. \cite{VS18} extends the results of \cite{VV02} to the case where $b$ is not symmetric. For the models in the article, \cite{VV02} is a sufficient reference.

@@

\section{Modal analysis and comparison of linear models}
Modal analysis is a important concept in the study of vibration of elastic bodies and also 
useful in the comparison of linear models. The following section is a short discussion of 
the basic concepts of modal analysis and how it enables the comparison of linear models.

@@

\subsection{Validity of series solution}
Consider the following question: When is the formal series solution valid for the vibration problem with initial values $u(0) = u_0$ and $u'(0) = u_1$? To answer this question, we again refer to the article \cite{CVV18}.

@@

\subsection{Comparison of models}
The results on modal analysis and the validity of the series solution enable the comparison of different linear models. The solution of the model problems can be expressed as a series solution of the eigenvalues and eigenvectors of the models. Therefore the models can be compared by comparing the eigenvalues and eigenvectors of the models.

@@

\section{Finite Element Method}
In the section, the Finite Element Method (FEM) is applied to the models to obtain an eigenvalue problem that can be solved numerically.

\subsection{Two-dimensional model}\label{ssec:2D_Model:FEM}
Consider a cantilever two-dimensional elastic body, with a rectangular cross-section.

\subsubsection{Reference configuration for rectangular cross-section}\label{sssec:2D_Model:RefConf}
Let $\{e_1,e_2\}$ be a right-handed orthonormal basis for $R^2$. Denote the elastic body by $\Omega \in R^2$, with $(0,0)$ as the point of reference. For a rectangular cross-section, the body $\Omega$ can be described as,
\begin{eqnarray*}
	\Omega = \left \{ x \in E_2 \ | \ 0 \leq x_1 \leq 1, \ -\frac{h}{2} \leq x_2 \leq \frac{h}{2} \right \},
\end{eqnarray*} where $\partial \Omega$ denotes the boundary of $\Omega$. The boundary $\partial \Omega$ can be divided into the four distinct lines (or edges) as follows: \label{sym:En}

@@

In this configuration, the body is clamped rigidly to the surface at $x_1 = 0$ denoted as $\Sigma$ and the body is free-hanging on the other boundaries denoted by $\Gamma$.

\subsubsection{Cantilever elastic body}
Consider a two-dimensional elastic body clamped rigidly to a surface at $x_1 = 0$. The body is free-hanging on the other boundaries. This is Problem 2D in section \ref{ssec:2D_Model}.

@@

\subsection{Weak variational form}
Define the inertia space $V$ as the closure of $T(\Omega)$ in $H := H^1(0,1)\times H^1(0,1)$. Denote $X = L^2(0,1)\times L^2(0,1)$. The inertia space is $W  = X$ with norm $||\cdot||_W = \sqrt{c(\cdot,\cdot)}$.

\subsubsection{Problem 2DWV}
Find a function $u$ such that $\forall \ t > 0$, $u \in V$ with $\partial_t^2 u \in W$ and
\begin{eqnarray*}
	c(u,v) + b(u,v) & = & (Q,v)
\end{eqnarray*}
for all $v\in V$.

@@

\subsection{Three-dimensional model} \label{ssec:3D_Model:FEM}
Consider a cantilever three-dimensional elastic body, with a rectangular cross-section.

\subsubsection*{Reference configuration for rectangular cross-section}
Let $\{e_1,e_2,e_3\}$ be a right-handed orthonormal basis for $R^3$. Denote the elastic body as $\Omega \in R^3$ with $(0,0,0)$ the point of reference. For a rectangular cross-section, the body $\Omega$ can be described as
\begin{eqnarray*}
	\Omega = \left\{ x \in R^3 \ | \ 0 \leq x_1 \leq 1, \ -\frac{h}{2} \leq x_2 \leq \frac{h}{2} , \ -\frac{b}{2} \leq x_3 \leq \frac{b}{2}\right \}
\end{eqnarray*}
Let $\partial \Omega$ denote the boundary of the body. Divide $\partial \Omega$ into the six distinct flat surfaces as follows:

@@

\subsection{Weak variational form}

>>>>

\title{Comparison of linear beam theories}

\author{Rudi du Toit\corref{cor1}}
\ead{u15000258@tuks.co.za}
\cortext[cor1]{Corresponding author}
\author{Prof N.F.J. van Rensburg}
\address{Department of Mathematics, University of Pretoria, Pretoria, South Africa}

@@

\begin{abstract}
Beams, while commonly used in engineering, are complex three-dimensional objects. Simplified models are often used in vibration problems due to their reduced complexity and computational demand. However, the accuracy of these models in practical applications is not always certain. This article extends previous research to examine a two-dimensional cantilever beam model's validity using modal analysis and the Finite Element Method (FEM), comparing its solutions with those of a three-dimensional model. Our findings reveal that the beam's shape is a crucial factor in the model's validity for practical applications. For the appropriate beam shapes, the two-dimensional model is a valid approximation of the three-dimensional model. 
\end{abstract}

\begin{keyword}
	Beams \sep Cantilever Beams \sep Two-Dimensional Models \sep Three-Dimensional Models \sep Vibration Analysis \sep Modal Analysis \sep Finite Element Method
\end{keyword}

@@

\section{Introduction}
In the article \cite{LVV09}, the authors investigate the validity of a cantilever Timoshenko beam model. Validity refers to how well the solution of the simplified model compare to a more complex higher dimensional model. The reference model in \cite{LVV09} is a two-dimensional cantilever beam model.

In this article, we build upon \cite{LVV09} by further examining the validity of a two-dimensional model for a cantilever beam, using a three-dimensional model as the reference. We follow the same methodology as \cite{LVV09}, employing modal analysis and the Finite Element Method (FEM) to compare the two models. This extention is motivated by the authors of \cite{LVV09}.

To start, the models are presented in a dimensionless variational form. We explain how the two-dimensional model is derived from the three-dimensional model as a special case of plane stress. The well-posedness of the two models is not proved in this article, but the necessary existence and uniqueness results from \cite{VV02} are mentioned.

Next we introduce the concept of modal analysis with a general vibration problem. Through modal analysis, we derive a formal series solution for the vibration issue, which is a linear combination of the problem's eigenvectors and eigenvalues. This approach allows us to compare the solutions of the two models by focusing on a comparison of their eigenvectors and eigenvalues.

We then apply the Finite Element Method (FEM) to the two models and obtain a formula for the models to numerically solve the eigenvalue problem. The article \cite{BV13} provides essential convergence results for the FEM in vibration problems. The textbook \cite{SF73} also offers insights into the convergence of eigenvalues and eigenfunctions in a vibration problem when using FEM.

Following \cite{LVV09}, by considering the mode shapes, we can match up the eigenvalues and compare them. In the comparison, we consider different shapes of beams that are useful in practical applications. We find that the shape of the beam significantly influences the validity of the two-dimensional model. The results of the comparison are presented in the form of tables and graphs.

\section{Linear mathematical models}
	\subsection{Three-dimensional theory}
		Consider a vector valued displacement function $u \in \Omega \subset R^3$.

@@

		\subsubsection*{Constitutive Equation}
			\textcolor{red}{************ VOORSTEL om the verkort ************}

			For an isotropic material, Hooke's law can be written in terms of Young's modulus $E$ and Poisson's ratio $\nu$. And if the principal stresses $\sigma_i$ are all non-zero, then Hooke's law can be written in the following form:
			\begin{eqnarray}
				T = \left( \frac{E}{1+\nu} \right)\mathcal{E} + \frac{\nu E}{(1+\nu)(1-2\nu)}\textrm{Tr}(\mathcal{E})I \label{eq:3D_Model:CE}.
			\end{eqnarray} Here $\mathcal{E}$ is the strain tensor and $T$ is the stress tensor.

			\textcolor{red}{************ VOORSTEL ************}

			For an isotropic material, Hooke's law can be written in terms of Young's modulus $E$ and Poisson's ratio $\nu$ as follows.

@@

		\subsubsection*{Dimensionless form} \label{ssec:3D_Model:DimensionlessForm}
			Suppose $\ell$ represents some notable dimension (e.g.the length) of the elastic body and $G$ the shear modulus of elasticity.

@@

			\textbf{Remark} The constants introduced are the same as those in \cite{LVV09}.
			These dimsensionless constants allow for the comparison of the higher dimensional models to the Timoshenko beam model.

@@

		Divide the boundary of $\Omega$ into two distinct parts, referred to as $\Sigma$ and $\Gamma$. The following will be considered a model problem for a three-dimensional elastic body (also for the two-dimensional elastic body).
		\begin{enumerate}
			\item[] The equations of motion \eqref{eq:3D_Model:EM-D} and \eqref{eq:3D_Model:divT-D}, and Hooke's law is satisfied in $\Omega$;
			\item[] The displacement $u = u_\Sigma$ is specified on
				$\Sigma$; The traction $Tn = t_\Gamma$ is specified on
				$\Gamma$.
		\end{enumerate} \label{sym:n}\label{sym:SigmaGamma}

		\subsubsection*{Cantilever model}
			Find a vector valued function $u$, satisfying equations
			\eqref{eq:3D_Model:EM-D} to \eqref{eq:3D_Model:CE-D} and the following boundary
			conditions:

			Boundary Conditions:
			\begin{eqnarray*}
				u & = & 0 \quad \textrm{ on } \Sigma\\
				Tn & = & 0 \quad \textrm{ on } \Gamma
			\end{eqnarray*}

			This will be referred to as \textbf{Problem 3D}.

		\subsection*{Variational form}\label{ssec:3D_Model:VariationalForm}
			Define the bilinear forms
			\begin{eqnarray}
				b(u,\phi) = \int_\Omega~c_1\textrm{Tr}({\cal E}\Phi)+ c_2\textrm{Tr}({\cal E})
				\textrm{Tr}(\Phi) ~dV \label{eq:3D_Model:Bilinear}
			\end{eqnarray}\label{sym:b}
			and
			\begin{eqnarray}
				c(u,\phi) = \int_\Omega~ (\partial^2_t u) \cdot \phi~dV \label{eq:3D_Model:Bilinear_c}
			\end{eqnarray}
			with $\displaystyle c_1 = \frac{1}{\gamma(1+\nu)}$ and $\displaystyle c_2 = \frac{\nu}{\gamma(1+\nu)(1-2\nu)}$.

@@

			Find a function $u$ such that for all $t>0$, $u \in T(\Omega)$ and
			\begin{align}
				c(u,\phi) = -b(u,\phi) + (Q,\phi), \label{eq:3D_Model:Problem3D1VEq}
			\end{align}
			for all $\phi \in T(\Omega)$

			This will be referred to as \textbf{Problem 3DV}.

	\subsection{Two-dimensional theory}
		\subsubsection*{General plane stress}
			Stresses acting parallel to a single plane are referred to as plane stress. 
			Consider the right-hand orthonormal set $\left\{e_1, e_2, e_3\right\}$. Without
			loss of generality, assume the stresses act parallel to the $e_1$-$e_2$ plane
			i.e. $\sigma_{3i} = 0$ for all $i$. From Hooke's Law \eqref{DM_H_E} the
			following equations are obtained: \label{sym:e_i}

@@

			This is known as general plane stress. The two-dimensional model is a special case of plane stress.

		\subsubsection*{The two dimensional model}
			Assume that $\sigma_{3i} = 0$ for $i = 1, 2, 3$; $\partial_3 u_1 = 0$, $\partial_3 u_2 = 0$ ($u_1$ and $u_2$ are functions of $x_1$ and $x_2$), and the strain component $\varepsilon_{33} = 0$. Then using the definition of strain 
			\begin{eqnarray}
				\partial_i u_3 = 0 \quad \textrm{ for } i = 1,2,3 \textrm{ on } \Omega.
			\end{eqnarray}

@@

			The equations of motion and constitutive equations follow:
			\subsubsection*{Equation of motion}\label{sssec:2D_Model:EquationOfMotion}
			\begin{eqnarray}
				\partial_t^2 u & = & \textrm{div}T + Q, \label{eq:2D_Model:EM}
			\end{eqnarray}
			where
			\begin{eqnarray}
				\textrm{div} T & = &
				\begin{bmatrix}
					\partial_1 \sigma_{11} + \partial_2 \sigma_{12} \\
					\partial_1 \sigma_{21} + \partial_2 \sigma_{22}
				\end{bmatrix}. \label{eq:2D_Model:DivT}
			\end{eqnarray}
			\subsubsection*{Constitutive equation}\label{sssec:2D_Model:ConstitutiveEquation}
			\begin{eqnarray}
				T = \frac{1}{\gamma(1+\nu)}\mathcal{E} + \frac{\nu}{\gamma(1-\nu^2)}\textrm{Tr}(\mathcal{E}) I. \label{eq:2D_Model:CE}
			\end{eqnarray}

		\subsubsection*{Cantilever model}

			Find a vector valued function $u$, satisfying equations \eqref{eq:2D_Model:EM} to \eqref{eq:2D_Model:CE} and the boundary conditions:

			Boundary Conditions:
			\begin{eqnarray*}
				u & = & 0 \quad \textrm{ on } \Sigma,\\
				Tn & = & 0 \quad \textrm{ on } \Gamma.
			\end{eqnarray*} with $n$ the outward unit normal.

			This will be referred to as \textbf{Problem 2D}.

		\subsubsection*{Variational form}\label{ssec:2D_Model:VariationalForm}
			Define the bilinear forms
			\begin{eqnarray}
				b(u,\phi) = \int_\Omega~c_1\textrm{Tr}({\cal E}\Phi)+ c_2\textrm{Tr}({\cal E})
				\textrm{Tr}(\Phi) ~dA \label{eq:2D_Model:Bilinear}
			\end{eqnarray}
			and
			\begin{eqnarray}
				c(u,\phi) = \int_\Omega~ (\partial^2_t u) \cdot \phi~dA \label{eq:2D_Model:Bilinear_c}
			\end{eqnarray}
			with $\displaystyle c_1 = \frac{1}{\gamma(1+\nu)}$ and $\displaystyle c_2 = \frac{\nu}{\gamma(1-\nu^2)}$.
			Also define the integral
			\begin{eqnarray}
				(f,g) &=& \int_{\Omega} f\cdot g \ dA \label{eq:2D_Model:Bilinear_int}
			\end{eqnarray}

			\subsubsection*{Test functions for Problem 2D}\label{sssec:2D_Model:TestFunctions2D1}
			Define the following set of test functions $T(\Omega)$ for Problem 2D:
			\begin{eqnarray*}
				T(\Omega) & = & \left\{ \phi \in C^1(\bar{\Omega}) \ | \ \phi = 0 \ \textrm{ on } \ \Gamma \right\}. 
			\end{eqnarray*}\label{sym:baromega}

			Find a function $u$ such that for all $t>0$, $u \in T(\Omega)$ and 
			\begin{align}
				c(u,\phi) = -b(u,\phi) + (Q,\phi) \label{eq:2D_Model:Problem2D1VEq}
			\end{align}
			for all $\phi \in T(\Omega)$. This will be referred to as \textbf{Problem 2DV}.

	\subsection{Timoshenko theory}
		Consider a Timoshenko beam with no damping defined on the interval $[0,\ell]$.
		Let $w$ represent the transverse displacement and $\phi$ a rotation of the cross-sections of the beam.

		\subsubsection*{Equations of motion}\label{sssec:1D_Model:EquationOfMotion}
			\begin{eqnarray}
				\rho A \partial_t^2 w &=& \partial_x V + Q, \label{eq:1D_Model:EquationOfMotion1}\\
				\rho I\partial_t^2 \phi  &=& V + \partial_x M. \label{eq:1D_Model:EquationOfMotion2}
			\end{eqnarray}\label{sym:A}\label{sym:V}\label{sym:Iinertia}\label{sym:M}\label{sym:phiT}
			In~\eqref{eq:1D_Model:EquationOfMotion1} and \eqref{eq:1D_Model:EquationOfMotion2} $\rho$ denotes the density, $A$ is the area of a cross section, $I$ is the area moment of inertia, $M$ is the moment, $V$ is the shear force and $Q$ is an external force acting on the beam.

		\subsubsection*{Constitutive equations}\label{sssec:1D_Model:ConstitutiveEquation}
			\begin{eqnarray}
				M &=& EI\partial_x \phi, \label{eq:1D_Model:ConstitutiveEquations1}\\
				V &=& AG\kappa^2(\partial_x w - \phi), \label{eq:1D_Model:ConstitutiveEquations2}\label{sym:kappa2T}
			\end{eqnarray}
			$E$ is Young's modulus, $G$ the shear modulus and $\kappa^2$ the shear correction factor.

		\subsubsection*{Dimensionless form}\label{sssec:1D_Model:DimensionlessForm}
			As mentioned in the dimensionless form of the three-dimensional model, the same dimensionless scaling can be applied to the Timoshenko beam theory.

			Set \[\tau = \frac{t}{t_0}, \,\, \xi = \frac{x}{\ell}, \,\, w^*(\xi,\tau) = \frac{w(x,t)}{\ell} \ \text{ and } \ \phi^*(\xi, \tau) = \phi(x,t).\]

			The dimensionless forms of the shear force, moment and external force density are \[ V^{*}(\xi,\tau) = \frac{V(x,t)}{AG\kappa^2}, \quad M^{*}(\xi,\tau) = \frac{M(x,t)}{A G\kappa^2 \ell} \,\,\, \ \text{and} \ \,\,\, Q^*(\xi,\tau) = \frac{Q(x,t)\ell}{A G\kappa^2}.\]

			Choose $t_0$ the same as the three-dimensional model, i.e. \[t_0 = \ell\sqrt{\frac{\rho}{G\kappa^2}}.\]

			As in \cite{LVV09} we use the dimensionless constants
			\begin{equation*}
				\alpha = \frac{A \ell^2}{I} \,\,\, \text{and} \,\,\, \beta
				=\frac{AG\kappa^2 \ell^2}{EI}.
			\end{equation*}\label{sym:alphabeta}

			Interestingly, it turns out that
			\begin{eqnarray*}
				\frac{\beta}{\alpha} & = & \gamma
			\end{eqnarray*} where $\gamma$ is the dimensionless parameter defined in the three-dimensional case.\\

		\subsubsection*{Dimensionless equations of motion}\label{sssec:1D_Model:DimensionlessEquationsOfMotion}
			\begin{eqnarray}
				\partial_{t}^{2} w &=& \partial_{x}V + Q, \label{eq:1D_Model:EquationOfMotion1D}\\
				\frac{1}{\alpha} \partial_{t}^{2} \phi &=& V + \partial_{x}M.\label{eq:1D_Model:EquationOfMotion2D}
			\end{eqnarray}
		\subsubsection*{Dimensionless constitutive equations}\label{sssec:1D_Model:DimensionlessConstitutiveEquations}
			\begin{eqnarray}
				M &=& \frac{1}{\beta}\partial_x \phi, \label{eq:1D_Model:ConstitutiveEquations1D}\\
				V &=& \partial_x w-\phi. \label{eq:1D_Model:ConstitutiveEquations2D}
			\end{eqnarray}
		The original notation is retained for convenience.

		\subsection{Models and variational forms}
			\subsubsection*{Bilinear forms and integral}
				For $f,g \in T[0,1]$, define the bilinear forms
				\begin{eqnarray}
					c(f,g) & = & \int_0^1  \partial_t^2 f_1 g_1 + \frac{1}{\alpha}\int_0^1  \partial_t^2 f_2g_2, \label{eq:1D_Model:Bilinear} \\
					b(f,g) & = &   \int_0^1 (f_1'-f_2)(g_1' - g_2) + \frac{1}{\beta}\int_0^1 f_2' g_2', \label{eq:1D_Model:Bilinear_c}
				\end{eqnarray}
				Define the integral
				\begin{eqnarray}
				(f,g)
					&=& \int_{0}^1 fg \label{eq:1D_Model:Bilinear_int}
				\end{eqnarray} for all $f,g \in L^2(0,1).$\\

				\textbf{Remark}\\
				$L^2(a,b)$ is the space of all square integrable functions on the interval $(a,b)$. The inner product is defined by $\displaystyle (f,g) = \int_a^b fg$, and the induced norm $\displaystyle ||f|| = \int_a^b f^2$.\\

			\subsection*{Pinned-pinned beam}
				Find functions $w$ and $\phi$ satisfying equations \eqref{eq:1D_Model:EquationOfMotion1D} to \eqref{eq:1D_Model:ConstitutiveEquations2D} and the boundary conditions:

				{Boundary Conditions}\\
				\begin{eqnarray*}
					w(0,\cdot) = 0, \ \ &M(0,\cdot) = 0, \label{eq:1D_Model:ProblemT1BC1}\\
					w(1,\cdot) = 0, \ \ &M(1,\cdot) = 0. \label{eq:1D_Model:ProblemT1BC2}
				\end{eqnarray*}

				This will be referred to as \textbf{Problem T-1}. Define the set of test functions $T[0,1]$ for Problem T-1 as
				\begin{eqnarray*}
					T[0,1] &:=&  \left\{f \in C^1[0,1] \ | \ f(0) = f(1) = 0 \right\} \times C^1[0,1]
				\end{eqnarray*}

				Then the variational problem for Problem T-1 is given by
				\subsubsection*{Problem T-1V}\label{sssec:1D_Model:ProblemT1V}
					Find a function ${u} = \langle w, \phi \rangle$ such that for all $t >0$, ${u} \in  T[0,1]$ satisfying
					\begin{eqnarray}
						c(\partial_t^2 u,{\phi}) &=& -b({u},{\phi}) + (Q,{\phi})
					\end{eqnarray} for each ${\phi} = \langle v, \psi \rangle \in T[0,1]$.

			\subsection*{Cantilever beam}
				Find functions $w$ and $\phi$ satisfying equations \eqref{eq:1D_Model:EquationOfMotion1D} to \eqref{eq:1D_Model:ConstitutiveEquations2D} and the boundary conditions:

				{Boundary Conditions}\\
				\begin{eqnarray*}
					w(0,\cdot) = 0, \ \ &\phi(0,\cdot) = 0, \label{eq:1D_Model:ProblemT2BC1}\\
					M(1,\cdot) = 0, \ \ &V(1,\cdot) = 0. \label{eq:1D_Model:ProblemT2BC2}
				\end{eqnarray*}

				This will be referred to as \textbf{Problem T-2}.  Define the set of test functions $T[0,1]$ for Problem T-2 as
				\begin{eqnarray*}
					T[0,1] &:=& \left\{g \in C^1[0,1] \ | \ g(0) = 0 \right\} \times \left\{g \in C^1[0,1] \ | \ g(0) = 0 \right\}
				\end{eqnarray*}

				Then the variational problem for Problem T-2 is given by
				\subsubsection*{Problem T-2V}\label{sssec:1D_Model:ProblemT1V}
					Find a function ${u} = \langle w, \phi \rangle$ such that for all $t >0$, ${u} \in  T[0,1]$ satisfying
					\begin{eqnarray}
						c(u,{\phi}) &=& -b({u},{\phi}) + (Q,{\phi}) \label{var_form_timo}
					\end{eqnarray} for each ${\phi} = \langle v, \psi \rangle \in T[0,1]$.

\section{Modal analysis and comparison of linear models}

	\subsection*{Existance and uniqueness of solutions}
		Our two models, the three-dimensional modela and the two-dimensional model, concern the vibration of elastic bodies. These type of models have a similar variational form as the wave equation and are called second order hyperbolic type problems. The articles \cite{VV02} and \cite{VS18} provide existence and uniqueness results for these type of problems. In \cite{VV02}, it is assumed that $b$ is symmetric. \cite{VS18} extends the results of \cite{VV02} to the case where $b$ is not symmetric. For the models in the article, \cite{VV02} is a sufficient reference.

@@

	\subsection*{Modal Analysis}
		Modal analysis is a important concept in the study of vibration of elastic bodies and also useful in the comparison of linear models. The following section is a short discussion of the basic concepts of modal analysis and how it enables the comparison of linear models.

@@

	\subsection*{Validity of series solution}
		Consider the following question: When is the formal series solution valid for the vibration problem with initial values $u(0) = u_0$ and $u'(0) = u_1$? To answer this question, we again refer to the article \cite{CVV18}.

@@

	\subsection*{Comparison of models}
		The results on modal analysis and the validity of the series solution enable the comparison of different linear models. The solution of the model problems can be expressed as a series solution of the eigenvalues and eigenvectors of the models. Therefore the models can be compared by comparing the eigenvalues and eigenvectors of the models.

@@

\section{Cantilever two-dimensional model}\label{ssec:2D_Model:FEM}
Let $\{e_1,e_2\}$ be a right-handed orthonormal basis for $R^2$. Denote the elastic body by $\Omega \in R^2$, with $(0,0)$ as the point of reference. For a rectangular cross-section, the body $\Omega$ can be described as,
\begin{eqnarray*}
	\Omega = \left \{ x \in E_2 \ | \ 0 \leq x_1 \leq 1, \ -\frac{h}{2} \leq x_2 \leq \frac{h}{2} \right \},
\end{eqnarray*} where $\partial \Omega$ denotes the boundary of $\Omega$. The boundary $\partial \Omega$ can be divided into the four distinct lines (or edges) as follows: \label{sym:En}

@@

In this configuration, the body is clamped rigidly to the surface at $x_1 = 0$ denoted as $\Sigma$ and the body is free-hanging on the other boundaries denoted by $\Gamma$. This is a catilever configuration.

@@

@@

\section{Cantilever three-dimensional model} \label{ssec:3D_Model:FEM}
Let $\{e_1,e_2,e_3\}$ be a right-handed orthonormal basis for $R^3$. Denote the elastic body as $\Omega \in R^3$ with $(0,0,0)$ the point of reference. For a rectangular cross-section, the body $\Omega$ can be described as
\begin{eqnarray*}
	\Omega = \left\{ x \in R^3 \ | \ 0 \leq x_1 \leq 1, \ -\frac{h}{2} \leq x_2 \leq \frac{h}{2} , \ -\frac{b}{2} \leq x_3 \leq \frac{b}{2}\right \}
\end{eqnarray*}
Let $\partial \Omega$ denote the boundary of the body. Divide $\partial \Omega$ into the six distinct flat surfaces as follows:

@@
